---
title: "Establishing adjacency between clusters"
author: "Lauren Mabe"
date: "4-14-2022"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Abstract
Correctly establishing adjacency between clusters is a critical component of the model to create efficient, compact clusters that do not “jump” over one another. There are many commonly used measures of adjacency using polygons, however these were found to be unsuitable for use with the FWG. Converting the point clusters into polygons using convex or concave hulls was considered as well as generating Voronoi polygons from the AD centroids. A new method was developed that measures the distance between the closest two FW points in a given pair of clusters. The adjacency function is used any time FW is exchanged between ADs and the adjacency matrix can be determined for a single AD at a time using minimal computation time. (the distance/adjacency matrix is really a distance/adjacency vector - 1D). Natural breaks in the distance vector are used to classify ADs (that are not the target AD) into adjacent and nonadjacent groups. 


# Setup
## Packages
```{r, warning = FALSE, message = FALSE}

# for spatial data
library(sp)
library(raster)

# for sample data built into package
library(spData)


```

## Data
A spatial points dataset from spData package is used as a substitute for the Food Waste Geography (FWG) dataset. The points represent commercial businesses that dispose of FW in Los Angeles County (BUS). Each point is randomly assigned a FW_dis value (tons of FW disposed of annually).
```{r}

set.seed(666) # for random number generation


# Data on 25,357 single family homes sold in Lucas County, Ohio, 1993-1998
# we are pretending they are commercial businesses, not houses
data(house)

# create a "FW_dis" column using random numbers from 0:5
# this will simulate the "FW Geography" dataset
house$FW_dis <- runif(nrow(house), 0, 5)

house$idx <- 1:nrow(house)

# drop the built-in columns
house <- house[,which(names(house) %in% c("idx","FW_dis"))]



plot(house, pch = 19, col = "black", main = "sample FWG")


```


## Generate inital clusters using K-means
The first step of the FWADA model is to use unsupervised K-means clustering to generate an initial set of clusters. The number of clusters to find is set as the minimum number of ADs needed at a given capacity to digest all FW in the FWG. In this example, AD capacity is set to 5000 tons/year, which will generate 13 clusters. The ADs are placed at the cluster centers.

```{r, message = FALSE}

# FW_kmeans() performs kmeans clustering and calculates FW_gen & FW_dis for each cluster
# @param Bus_pnts - business points w/ assigned FW in the cluster
# @param AD_cap - the capacity of the ADs to be placed
# Returns a named list ("object" type I call a kmeans_result/km_res)
# [1] BUS - the business cluster (spdf)
# [2] AD - cluster centroids representing the AD (spdf), this section contains the data about the cluster
FW_kmeans <- function(Bus_pnts, AD_cap) {
    
    
    # calculate number of ADs needed
    # rounded up, the 1.0 means we are capturing 100% of FW
    cent <- ceiling((sum(Bus_pnts$FW_dis) * 1.0) / AD_cap)
    
    message(paste0(cent, " clusters"))
    
    # Convert to matrix
    mBUS <- as.matrix(Bus_pnts@coords)
    
    # kmeans from stats package
    kmcluster <- kmeans(mBUS, centers = cent, iter.max = 500, nstart = 1)
    
    # Create df to add coords to
    AD <- data.frame(AD_code = paste0("x",1:nrow(kmcluster$centers)))
    coordinates(AD) <- kmcluster$centers
    
    AD@proj4string <- Bus_pnts@proj4string
    
    AD$initial_K <- cent
    AD$AD_cap <- AD_cap
    
    # Add cluster group to each BUS point
    BUS <- "empty"
    BUS <- cbind(Bus_pnts, paste0("x",as.factor(kmcluster$cluster))) 
    colnames(BUS@data)[ncol(BUS@data)] <- c("AD_code")
    
    # calculate FW in each cluster
    AD$FW_gen <- sapply(AD$AD_code, function(x) sum(BUS$FW_gen[BUS$AD_code == x]))
    AD$FW_dis <- sapply(AD$AD_code, function(x) sum(BUS$FW_dis[BUS$AD_code == x]))
    
    ret <- list(AD, BUS)
    names(ret) <- c("AD", "BUS")
    
    
    return(ret)
} #close FW_kmeans()




set.seed(666) # repeatable

# get clusters using FW_kmeans function
## returns a named list containing FW points w/ assigned AD_code (BUS) 
## and their centroids/the AD with information about the cluster (AD)
km_house <- FW_kmeans(house, 5000)


# plot the result
plot(km_house$BUS, col = as.factor(km_house$BUS$AD_code), pch = 20, 
     main = "Kmeans clustering of sample data")
points(km_house$AD, pch = 24, col = "white", bg = "black")

```


# Distance Matrix 
The distance matrix used to establish adjacency is based on the boundary of the clusters instead of centroid based methods. While the boundary of the clusters can be drawn using convex or concave hulls, these methods are computationally complex and take time to solve. To save memory, the distance between the pair of FW points that are closest in each cluster is measured for the distance matrix. Since adjacency is established for a single AD at a time, this matrix is really just a distance vector. In this section, the AD that needs to exchange FW points with its neighbors is shown in red (AD code: x1) and is referred to as the target cluster or AD.

## Closest two points between clusters  
First, the point in the target cluster (x1) that is closest to the AD of another cluster is found and visa verse. Then the distance between those two FW points is measured for the distance vector. This is done iteratively for all clusters until the full distance vector for the target AD (x1) is found. The length of the red lines in the following plot is the distance vector for the target AD.

```{r, echo = FALSE, results='hide',fig.keep='all'}

# plots the closest pair of points from on AD (OG - original) to another (T - target)
# @param OG_code - Default "x1", the AD for for the OG AD (the one finding adjacency for)
# @param T_code - the AD_code for the Target AD.
# @param km_res - a kmeans result. A named list (AD = AD, BUS = BUS)
# returns plots of the closest points
plot_closest_points <- function(OG_code = "x1", T_code, km_res) {
    
    # pull out the BUS points from km_res
    OG_BUS <- km_res$BUS[km_res$BUS$AD_code == OG_code,]
    T_BUS <- km_res$BUS[km_res$BUS$AD_code == T_code,]
    
    # pull out the AD points from km_res
    OG_AD <- km_res$AD[km_res$AD$AD_code == OG_code,]
    T_AD <- km_res$AD[km_res$AD$AD_code == T_code,]
    
    # point distance from OG_BUS to to T_AD
    OG_BUS$pdist <- pointDistance(OG_BUS, T_AD, latlong = T)
    
    # pnt dist from T_BUS to OG_AD
    T_BUS$pdist <- pointDistance(T_BUS, OG_AD, latlong = T)
    
    # pnt distance between the two points closest to opposite AD
    # this is the distance reported in the adjacency matrix
    dist2points <- pointDistance(OG_BUS[OG_BUS$pdist == min(OG_BUS$pdist),],
                                    T_BUS[T_BUS$pdist == min(T_BUS$pdist),])
    
    
    l <- as(rbind(OG_BUS[OG_BUS$pdist == min(OG_BUS$pdist),],
                  T_BUS[T_BUS$pdist == min(T_BUS$pdist),]), "SpatialLines")
    # plot the result
    plot(km_res$BUS, col = "grey",
         main = paste0("x1 to ", T_code, " | dist: ", round(dist2points))) # all points
    points(OG_BUS, pch = 19, col = "red") 
    points(T_BUS, pch = 19, col = "blue")  
    points(OG_AD, pch = 24, col = "white", bg = "black") 
    points(T_AD, pch = 24, col = "white", bg = "black")  
    points(OG_BUS[OG_BUS$pdist == min(OG_BUS$pdist),], pch = 15, col = "black")
    points(T_BUS[T_BUS$pdist == min(T_BUS$pdist),],pch = 15, col = "black")
    plot(l, add = TRUE, lwd = 2, col = "red")
}

par(mfrow = c(3,3), mar = c(1,1,1,1))

# lapply over the AD codes to get distance of closest points from x1 to all other ADs
temp <- lapply(km_house$AD$AD_code, plot_closest_points, km_res = km_house, OG_code = "x1")
temp
```



## Ordered distance vector
getDistMatRow() is the function developed to calculate the distance vector. It also puts the vector into ascending order.

```{r}

# calculates the adjacency matrix row for a given AD_code
# @param AD_code - AD to get adjacents for
# @param kmean_res - the ADs and BUSs
# returns a vector, the ordered distance/adjacency matrix for a given AD
getDistMatRow <- function(AD_code, km_res) {
    
    # plots the closest pair of points from on AD (OG - original) to another (T - target)
    # @param OG_code - Default "x1", the AD for for the OG AD (the one finding adjacency for)
    # @param T_code - the AD_code for the Target AD.
    # @param km_res - a kmeans result. A named list (AD = AD, BUS = BUS)
    # returns - the distance between closest BUS points (numeric)
    simpleClosePoints <- function(OG_code, T_code, km_res) {
    
        # get ADs
        OG_AD <- km_res$AD[km_res$AD$AD_code == OG_code,]
        T_AD <- km_res$AD[km_res$AD$AD_code == T_code,]
        
        # get BUS
        OG_BUS <- km_res$BUS[km_res$BUS$AD_code == OG_code,]
        T_BUS <- km_res$BUS[km_res$BUS$AD_code == T_code,]
        
        # point distance from OG_BUS to to T_AD
        OG_BUS$pdist <- pointDistance(OG_BUS, T_AD, lonlat = FALSE)
        
        # pnt dist from T_BUS to OG_AD
        T_BUS$pdist <- pointDistance(T_BUS, OG_AD, lonlat = FALSE)
        
        # closest BUS point to the other AD
        p1 <- OG_BUS[OG_BUS$pdist == min(OG_BUS$pdist, na.rm = TRUE),]
        p <- T_BUS[T_BUS$pdist == min(T_BUS$pdist, na.rm = TRUE),]

        # pnt dist from two closest BUS points
        ret <- pointDistance(p1, p, lonlat = FALSE)
        
        # return the distance between the closest two points
        return(ret)
    } # close simpleClosePoints()
    
    # list of AD codes
    code_list <- as.character(km_res$AD$AD_code)
    
    
    # loop through list of AD codes,
    # find dist from closest BUS points to T_code (the code being updated for)
    # there will be a 0 in this vec. In a dist_mat, this 0 would be the diagonal
    dist_vec <- sapply(code_list, simpleClosePoints, T_code = AD_code, km_res = km_res)
    names(dist_vec) <- code_list
    return(dist_vec)
    
} # close getDistMatRow()

        
# get the sorted dist_mat row for the AD_code we need
sorted <- sort(getDistMatRow("x1", km_house), FALSE)

# remove the 0 from the dist_vec (its been sorted to the front).
# This is the diagonal on a distance matrix and is x1 to x1
sorted <- sorted[2:length(sorted)]

sorted
```

When the distance vector is put in ascending order and plotted, groups of distances are observed. Since the distance vector is based on the points nearest to the edge of the cluster and not the AD at the centroid, the height of these groups on the Y-axis is approximately the diameter of the clusters. These groups on the plot therefore correspond to rings of clusters around the original AD (x1)

```{r, echo = FALSE}

par(mfrow = c(1,2), mar = c(3,4,1,1))
    
plot(sorted, pch = 15, col = "black",
     ylab = "distance between closest points",
     main = "Ordered adjacency matrix")

plot(km_house$BUS, pch = 20, col = as.factor(km_house$BUS$AD_code), 
     main = "Map of clusters")
points(km_house$BUS[km_house$BUS$AD_code == "x1",], pch = 20, col = "red")
points(km_house$AD, pch = 24, col = "white", bg = "black")
legend("bottomright", pch = 19, col = "red", legend = c("AD_code x1"))



```

# Adjacency Matrix

## Identify Natural Breaks
To get classify clusters into adjacent and nonadjacent groups, the breaks in this vector need to be established. Jenk's Natural Breaks algorithm was considered, as it is an unsupervised method to cluster one-dimensional objects. Much like K-means clustering, it requires a user-specified number of clusters to find as an input. Given that the FWADA model is run at multiple AD capacities, we were unable to determine an appropriate value. 

To generate efficient, compact clusters, FW should only be passed between clusters immediately surrounding the target AD (x1). Therefore, it is only necessary to identify the first break in the ordered distance matrix. To do so, the difference between points in the ordered distance vector is calculated (middle plot). The first natural break is identified where this difference is much larger than the others. After parameter testing, this value was set as a distance above 90% of standard deviation, shown in blue on the middle plot. Clusters with distance vector below this break were considered adjacent to the target AD.

```{r}

# make a df of the distance between each point of the sorted adjacency matrix
diff_df <- data.frame(diff = diff(sorted),
                      node = 2:length(sorted))

# find standard dev of these differences
sd_diff <- sd(diff_df$diff)

# get points above 90% of the standard deviation
diff_df$above_SD <- FALSE
diff_df$above_SD[diff_df$diff >= (sd_diff * 0.9)] <- TRUE

# get the index of the first point above standard dev line
# the points with distances below the cutoff point this will be the adjacents
# ex: cutoff = 7, then first 6 points (based on dist mat) will be adjacents - this is why we subtract one
cutoff <- min(diff_df$node[diff_df$above_SD == TRUE], na.rm = TRUE)
cutoff <- cutoff - 1

# get the AD_codes that will be adjacent
adj_ADs <- names(sorted[1:cutoff])

adj_ADs

```

```{r, echo = FALSE}

# plot results
par(mfrow = c(1,3), mar = c(3,4,1,1))

## Left
plot(sorted, pch = 15, col = "black",
     ylab = "distance between closest points",
     main = "Ordered adjacency matrix")
points(sorted[1:4], pch = 15, col = "blue")

## middle
plot(diff_df$node, diff_df$diff, pch = 15, col = "black", 
     main = "difference plot", 
     ylab = "difference between sorted adj. mat. distances")
lines(diff_df$node, diff_df$diff)
abline(h = sd_diff, col = "blue")

## right
plot(km_house$BUS, pch = 20, col = "grey", 
     main = "adjacent clusters")
points(km_house$BUS[km_house$BUS$AD_code == "x1",], pch = 20, col = "red")
points(km_house$BUS[km_house$BUS$AD_code %in% adj_ADs,], col = "blue")
points(km_house$AD, pch = 24, col = "white", bg = "black")
legend("bottomright", pch = 19, col = "red", legend = c("AD_code x1"))



```

