---
title: "Functions to equalize FW between clusters"
author: "Lauren Mabe"
date: "4-13-2022"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract
This file explains the two spatial processes the FWADA model uses to manipulate clusters in order to equalize FW between ADs. Two functions were developed to manipulate clusters and change the amount of FW allocated to an AD. The splitting function splits large clusters in half while the grabbing function exchanges FW between adjacent clusters.

# Setup
## Load libraries
```{r, warning = FALSE, message = FALSE}
# for spatial data
library(sp)
library(raster)

# for sample data built into package
library(spData)

# for visualizing principal direction
library(ggfortify)


```


## Load data
A spatial points dataset from spData package is used as a substitute for the Food Waste Geography (FWG) dataset. The points represent commercial businesses that dispose of FW in Los Angeles County (BUS). Each point is randomly assigned a FW_dis value (tons of FW disposed of annually). 
```{r}

set.seed(666) # for random number generation


# Data on 25,357 single family homes sold in Lucas County, Ohio, 1993-1998
data(house)

# create a "FW_dis" column using random numbers from 0:5
  # this will simulate the "FW Geography" dataset
house$FW_dis <- runif(nrow(house), 0, 5)

# index column
house$idx <- 1:nrow(house)

# drop the built-in columns
house <- house[,which(names(house) %in% c("idx","FW_dis"))]



plot(house, pch = 19, col = "black", main = "sample FWG")


```

## Generate inital clusters using K-means
The first step of the FWADA model is to use unsupervised K-means clustering to generate an initial set of clusters. The number of clusters to find is set as the minimum number of ADs needed at a given capacity to digest all FW in the FWG. In this example, AD capacity is set to 5000 tons/year, which will generate 13 clusters. The ADs are placed at the cluster centers. 
```{r, message = FALSE}

#~~~~~~~~~~~~~~~FW_kmeans() function ~~~~~~~~~~~~~~~~~~
# FW_kmeans() performs kmeans clustering and calculates FW_gen & FW_dis for each cluster
# @param Bus_pnts - business points w/ assigned FW in the cluster
# @param AD_cap - the capacity of the ADs to be placed
# Returns a named list ("object" type I call a kmeans_result/km_res)
# [1] BUS - the business cluster (spdf)
# [2] AD - cluster centroids representing the AD (spdf), this section contains the data about the cluster
FW_kmeans <- function(Bus_pnts, AD_cap) {
    
    
    # calculate number of ADs needed
    # rounded up, the 1.0 means we are capturing 100% of FW
    cent <- ceiling((sum(Bus_pnts$FW_dis) * 1.0) / AD_cap)
    
    message(paste0(cent, " clusters"))
    
    # Convert to matrix
    mBUS <- as.matrix(Bus_pnts@coords)
    
    # kmeans from stats package
    kmcluster <- kmeans(mBUS, centers = cent, iter.max = 500, nstart = 1)
    
    # Create df to add coords to
    AD <- data.frame(AD_code = paste0("x",1:nrow(kmcluster$centers)))
    coordinates(AD) <- kmcluster$centers
    
    AD@proj4string <- Bus_pnts@proj4string
    
    AD$initial_K <- cent
    AD$AD_cap <- AD_cap
    
    # Add cluster group to each BUS point
    BUS <- "empty"
    BUS <- cbind(Bus_pnts, paste0("x",as.factor(kmcluster$cluster))) 
    colnames(BUS@data)[ncol(BUS@data)] <- c("AD_code")
    
    # calculate FW in each cluster
    AD$FW_gen <- sapply(AD$AD_code, function(x) sum(BUS$FW_gen[BUS$AD_code == x]))
    AD$FW_dis <- sapply(AD$AD_code, function(x) sum(BUS$FW_dis[BUS$AD_code == x]))
    
    ret <- list(AD, BUS)
    names(ret) <- c("AD", "BUS")
    
    
    return(ret)
} #close FW_kmeans()


#~~~~~~~~~~~~~~~~~~~~~~~ Generate clusters ~~~~~~~~~~~~~~~~~~~~~~~~~`

set.seed(666) # repeatable

# get clusters using FW_kmeans function
## returns a named list containing FW points w/ assigned AD_code (BUS) 
## and their centroids/the AD with information about the cluster (AD)
km_house <- FW_kmeans(house, 5000)


# plot the result
plot(km_house$BUS, col = as.factor(km_house$BUS$AD_code), pch = 20, 
     main = "Kmeans clustering of sample data")
points(km_house$AD, pch = 24, col = "white", bg = "black")

```

# Splitting large clusters
If a cluster is over 140% full, it can be split equally with both new clusters above the 70% minimum operating capacity of the AD. Clusters are split perpendicularly to the principal direction through the weighted center using Principal Component Analysis (PCA). The eigenvectors returned by PCA and the bounding box of the cluster are used to generate polygons to split the cluster using the over() function. If the split through the middle of the cluster does not create relatively equal sized clusters, it will be moved incrementally towards the larger cluster until both clusters are equal (within 5% of the total FW within the cluster form one another). The ADs are located at the weighted center of the two new clusters. This function adds surplus AD capacity to the map which may be used to absorb FW from other clusters in later iterations of the FWADA model.


## Sample cluster
```{r, echo = FALSE}


# AD code of largest cluster by FW
cluster <- km_house$AD$AD_code[km_house$AD$FW_dis == max(km_house$AD$FW_dis)]

# get BUS points
cluster <- km_house$BUS[km_house$BUS$AD_code == cluster,]

plot(cluster, pch = 20, col = "black",
     main = paste0("Sample Cluster: ", cluster$AD_code[1], "| FW_dis: ", round(sum(cluster$FW_dis))))




```

## PCA for Principal Direction
Principal Component Analysis (PCA) is used on the cluster coordinates to determine the principal direction of the cluster. The eigenvectors are used to determine the line on which to split the clusters.
```{r}

# get the target FW before splitting the cluster, will need later
target_FW <- sum(cluster$FW_dis) / 2

# run principal direction on the cluster coordinates
pc <- princomp(cluster@coords)


# get the loadings for the two vectors 
pc_mat <- matrix(as.numeric(pc$loadings), ncol = 2)


# row/col names
colnames(pc_mat) <- c("x", "y")

# loadings matrix for the two eigenvectors
pc_mat


# plot the cluster/eigenvectors
autoplot(object = pc, loadings = TRUE,  main = "PCA (cluster coordinates)")


```

The angle of the eigenvectors gives the information about principal direction.  
NOTE: this plot has been turned 90 degrees clockwise for some reason....

### Calculate angles of loading matrix
The angle from the X axis (from East) is calculated for the principal direction (PD) (largest eigenvector). Other angles are calculated by adding 90 degrees to this value. The weighted center of the cluster and the bounding box coordinates are also located here.  
```{r}

#~~~~~~~~~~~~~~~~~~~ determine PD angles ~~~~~~~~~~~~~~~~~~~~

# get theta function. Returns theta when given the loadings matrix
get_theta <- function(i) {
    
  rad2deg <- function(rad) {(rad * 180) / (pi)} 
   
  t <- atan2(i["y"], i["x"]) # find angle from x axis
  t <- rad2deg(t) # convert to degrees

  return(t)
} # close get_theta()

# get the angle from x axis of the PD
  # PD is the eigenvector with max standard dev
theta <- get_theta(pc_mat[which(pc$sdev == max(pc$sdev)),])

# get the other angles by adding 90 degrees
theta <- c(theta, theta + 90, theta + 180, theta + 270)

# if the principle direction is in Q2 (Northwest), then the 4th theta (one that would be in Q1) is over 360
# subtract 360 from it to get angles between 0-360 degrees
if (theta[1] > 90) {
  theta[4] <- theta[4] - 360
}
  

#~~~~~~~~~~~~~~~~~ Weighted Center of cluster (AD location) ~~~~~~~~~~~~~
weightedCenter <- function(cluster1) {
  # get weighted center
  wcx <- weighted.mean(x = cluster1@coords[,1], 
                   w = cluster1$FW_dis)
  
  wcy <- weighted.mean(x = cluster1@coords[,2], 
                   w = cluster1$FW_dis)
  return(c(wcx, wcy))
} # close weightedCenter

wc <- weightedCenter(cluster)
  
  

#~~~~~~~~~~~~Bounding box of cluster ~~~~~~~~~~~~~~
  
# get bounding box
bb <- cluster@bbox

# change rownames to remember which one is which
rownames(bb) <- c("EW", "NS")



# bbox has the coordinates of the corners of the bbox
# the bbox is used as a boundary for the split polygon that splits the cluster
bb
```

```{r, echo = FALSE}

# make an sp object of bbox for plotting
bbox_sp <- as.matrix(rbind(c(bb[1,1], bb[2,2]), # NW
                           c(bb[1,2], bb[2,2]), # NE
                           c(bb[1,2], bb[2,1]), # SE
                           c(bb[1,1], bb[2,1]))) # SW
colnames(bbox_sp) <- c("x_coord", "y_coord")

# polygon for plotting bbox
bbox_sp <- Polygon(bbox_sp)
bbox_sp <- Polygons(list(bbox_sp), 1)
bbox_sp <- SpatialPolygons(list(bbox_sp))
bbox_sp@proj4string <- house@proj4string



# plot cluster, wc, bbox, and note the eigenvector angles
plot(cluster, pch = 20, col = "black", 
     main = "Sample Cluster w/ Weighted Center and bbox")
points(wc[1], wc[2], pch = 24, col = "black", bg = "red")
plot(bbox_sp, add = TRUE)
legend("bottomleft", legend = round(theta), title = "Calculated PD angles")


```

### Establish eigenvector/bbox intersections
Now that we have the angles of the eigenvectors, the weighted center, and the bbox, we can begin to split the cluster. To split the cluster, a polygon that covers half the cluster is generated to use with the over() (spatial join) function. To draw the splitting polygon, we first establish the where the eigenvector angles intersect the bbox as these will be used as corner points for drawing the split_poly (polygon used with over())
```{r}


#~~~~~~~~~~` Get polygon corner coordinates ~~~~~~~~~~~~
# function to get the PD point coords on the bbox for a single eigenvector angle at a time
# @param theta - the eigenvector angle to get the PD coord for. 
# @param bb1 - the bbox matrix
# Returns point coordinates for the split poly corners
getPDpointCoords <- function(theta, bb1 = bb) {
        
    # the quadrant the eigenvector angle is in determines what coordinates are chosen for x&y (taken from bbox)
    # trigonometry is used to determine where on the bbox to put the PD point in relation to the (weighted) center
    
    if (theta > 45 && theta < 135) {
        # on north side
        
        # find length of x
        # adj = opp/tan(theta)
        # length of opp (y) is distance from weighted center to N bbox
        d <- (bb1["NS", "max"] - wc[2]) / tan(theta*pi/180)
        
        # add/subtract that from center's x coord - this becomes the x coord
        x_coord <- wc[1] + d
        
        # y coord is north bbox y coord
        y_coord <- bb1["NS", "max"]
        
        # return coords
        return(c(x_coord, y_coord))
        
    } else if (theta > 135 && theta < 225) {
        # on east
        
        # change theta to interior angle
        theta <- 180 - theta
        
        # find length of y
        d <- (bb1["EW", "min"] - wc[1]) * tan(theta*pi/180)
        
        # add/sub that from centers y coord - this becomes the y coord
        y_coord <- wc[2] - d 
        
        # x coord is bbox x coord
        x_coord <- bb1["EW", "min"]
        
        # return coords
        return(c(x_coord, y_coord))
        
    } else if (theta > 225 && theta < 315) {
        # on south bbox
        
        # change theta to interior angle
        theta <- 270 - theta
        
        # find length of x
        d <- (bb1["NS", "min"] - wc[2]) * tan(theta*pi/180)
        
        # add/sub from center x coord - this becomes the x coord
        x_coord <- wc[1] + d 
        
        # y coord is bbox's y coord
        y_coord <- bb1["NS", "min"]
        
        # return coords
        return(c(x_coord, y_coord))
        
    } else if (theta > 315 || theta < 45) {
        
        # find length of y
        d <- (bb1["EW", "max"] - wc[1]) * tan(theta*pi/180)
        
        # add/sub from center y coord - this becomes the y coord
        y_coord <- wc[2] + d
        
        # x coord is bbox's x coord
        x_coord <- bb1["EW", "max"]
        
        # return coords
        return(c(x_coord, y_coord))
        
    } # close ifelse
    
} # close getPDpointCoords

# get the PD points for each theta using lapply
coord_mat <- do.call(rbind, lapply(theta, getPDpointCoords))
rownames(coord_mat) <- round(theta) # not super needed, just helps me keep things straight

#~~~~~~~~~~~~~~~ create splitting polygon ~~~~~~~~~~~~~~~~~~~~
# create SP object with these points
  # create dataframe to go with the points
coord_df <- data.frame(OID = 1:4,
                       theta = round(theta),
                       type = c("PD", "opp", "PD", "opp"),
                       PC_sdev = c(max(pc$sdev), min(pc$sdev), max(pc$sdev), min(pc$sdev)))

# create sp object
pd <- SpatialPointsDataFrame(coords = coord_mat, data = coord_df, proj4string = house@proj4string)

# plot PD points
plot(cluster, pch = 20, col = "black", 
     main = "cluster w/ PD points")
points(wc[1], wc[2], pch = 24, col = "black", bg = "red")
plot(bbox_sp, add = TRUE)
points(pd, pch = 19, col = "black")
points(pd, pch = 20, col = c("red", "blue", "yellow", "green"))
legend("bottomleft", pch = 20, col = c("red", "blue", "yellow", "green"),
       legend = round(theta), title = "PD angles")


```

## Split the cluster
The cluster is split by building the split_poly using 5 of the points established above. These points include 2 bbox corners, the weighted center, and the endpoints of the line perpendicular to the principal direction. FW generator points within the split_poly are reclassified to a new AD by adding "-1" to the AD_code and those outside are given "-2".
```{r}

#~~~~~~~~~~~~~~~ establish which line is the split line ~~~~~~~~~~~~~~~~ #
# this is determined by the distance between the two points, get split on shorter one
pd$length <- 999 #dummyvar

pd$length[pd$type == "PD"] <- pointDistance(pd[1,], pd[3,], lonlat = FALSE)
pd$length[pd$type == "opp"] <- pointDistance(pd[2,], pd[4,], lonlat = FALSE)


# designate the shortest line as the split line
pd$type <- as.character(pd$type)
pd$type[pd$length == min(pd$length)] <- "split"


# determine distance to weighted center
pd$dist_to_wc <- pointDistance(wc, pd, lonlat = FALSE)

#~~~~~~~~~~~~~~~~~~ form splitting polygon ~~~~~~~~~~~~~~~~~~~~~~~~~~ #
# createSplitPoly() creates the polygon to use to split the cluster
# @param pd_pnts - sp object representing the PD points of the cluster
# @param bb_pnts - an sp object representing the bbox corners
# returns a list
# $split_poly - the polygon used to split the cluster w/ over()
# $poly_points - points at the corners of the polygon, this holds the data for which point is which
createSplitPoly <- function(pd_pnts, bb_pnts) {
    
    split_line <- pd_pnts[pd_pnts$type == "split",] # the line designated as split
    
    
    # add the PD_point (point on non-split line furthest from weighted center)
    temp <- pd_pnts[pd_pnts$length == max(pd_pnts$length),] # the non-split line (has PD point as one of its ends)
    split_line <- rbind(split_line, temp[temp$dist_to_wc == max(temp$dist_to_wc),]) # the split line + PD point
    
    # add the bbox points where the coordinates of bbox_sp match the PD_point
    # https://stackoverflow.com/questions/28233561/finding-rows-containing-a-value-or-values-in-any-column
    # the apply goes over the rows of bbox_corner matrix, checks if any() of the values match the split line
    ## (basically, checks which side of split line the PD point is on)
    # its within a which() so arr.ind = TRUE returns the row indexes that match
    # this is within selecting from bbox_sp using [row,col] - selecting the rows w/ index returned from which()
    ## (selects the bbox corners that are on the same side as the PD point)
    poly_points <- rbind(split_line[,1:3],
                         bb_pnts[which(apply(bb_pnts@coords, 1,
                                             function(r) any(r %in% split_line[3,]@coords)) == TRUE, 
                                       arr.ind = TRUE),])
    
    # poly_points contains the split line endpoints, the PD point, and the bbox corners that go w/ PD point.
    ## the points need to be in the right order. need to go in a circle
    poly_points$OID <- 1:5
    poly_points <- poly_points[order(poly_points$theta),]
    
    
    # create the polygon
    # create a polygon with opp_pd box we made
    split_poly <- Polygon(poly_points)
    split_poly <- Polygons(list(split_poly),1)
    split_poly <- SpatialPolygons(list(split_poly))
    proj4string(split_poly) <- house@proj4string
    
    ret <- list(split_poly = split_poly, poly_points = poly_points)
    
    return(ret)
} # close createSplitPoly()



# create bbox corners sp object
bbox_df <- data.frame(OID = 1:4, 
                      theta = c(135, 225, 315, 45),
                      type = "bbox")
bbox_coords <- matrix(data = c(bb["EW", "min"], bb["NS", "max"],  
                               bb["EW", "min"], bb["NS", "min"], 
                               bb["EW", "max"], bb["NS", "min"], 
                               bb["EW", "max"], bb["NS", "max"]), ncol = 2, byrow = TRUE)
bbox_corners <- SpatialPointsDataFrame(coords = bbox_coords, data = bbox_df, proj4string = house@proj4string)





# split the polygon using the pd and bbox_sp points
split_poly <- createSplitPoly(pd, bbox_corners)

# separate return- poly points holds the dataframe we need for this
poly_points <- split_poly$poly_points
split_poly <- split_poly$split_poly

#~~~~~~~~~~~~~~~~~~~~~~~~~~ split the cluster using the split_poly ~~~~~~~~~~~~~~~~~~~~~~~~~~# 

# get what BUS points are in the split_poly and which are outside
cluster$split <- over(cluster, split_poly)

# get the centers of the two new clusters
wc1 <- weightedCenter(cluster[is.na(cluster$split) == TRUE,])
wc2 <- weightedCenter(cluster[is.na(cluster$split) == FALSE,])

# calculate FW in new clusters
check_df <- data.frame(x = c(wc1[1], wc2[1]),
                       y = c(wc1[2], wc2[2]),
                       FW_dis = c(sum(cluster$FW_dis[is.na(cluster$split) == TRUE]),
                                  sum(cluster$FW_dis[is.na(cluster$split) == FALSE])))


# plot result
plot(bbox_sp, lwd = 2, main = "Split Cluster")
points(cluster[is.na(cluster$split) == TRUE,], pch = 20, col = "red")
points(cluster[is.na(cluster$split) == FALSE,], pch = 20, col = "blue")
plot(split_poly, add = TRUE)
points(wc1[1], wc1[2], pch = 24, col = "black", bg = "white")
points(wc2[1], wc2[2], pch = 24, col = "black", bg = "white")
legend("bottomleft", pch = 19, col = c("blue", "red"),
       legend = c(paste0("x7-1: ", round((check_df$FW_dis[2]/5000)*100, 2), "% full"), 
                  paste0("x7-2: ", round((check_df$FW_dis[1]/5000)*100, 2), "% full")))


```

When running the principal direction split function in the FWADA model, it is done in a repeat loop. Within the loop, the function checks that the FW has been split relatively evenly (w/in 5% of total cluster FW). If it has, the loop breaks and the model moves on to the next cluster. If the cluster has not been split evenly, the function will move the split line towards the direction of the larger cluster at a user-specified incremental amount (given as a parameter to the model). In this case, the line needs to be moved towards the red cluster. This is line is moved until the cluster is split evenly within the repeat loop. There are some edge cases where the distance to move the line is too large and too much FW shifts between the two clusters. In this case, the splitting function will cut the distance to move the line in half so less FW moves between the two so as not to get stuck in a loop (not shown here).

```{r, echo = FALSE}

# coordinates of split poly corners
split_coords <- split_poly@polygons[[1]]@Polygons[[1]]@coords

# move all the Ys up (north) 500 feet
split_coords[,2] <- split_coords[,2] - 500

# the real function in the model uses the new split_coords to change the split_poly and resplit the cluster
# here, we just create a spatial lines object to plot
l <- SpatialLines(list(Lines(list(Line(rbind(split_coords[3,], 
                                             split_coords[2,]))), 1)))


# plot
plot(bbox_sp, main = "moving split line 500m north")
points(cluster[is.na(cluster$split) == TRUE,], pch = 20, col = "red")
points(cluster[is.na(cluster$split) == FALSE,], pch = 20, col = "blue")
plot(split_poly, lwd = 2, add = TRUE)
plot(l, add = TRUE, lty = 3, col = "black")
points(wc1[1], wc1[2], pch = 24, col = "black", bg = "white")
points(wc2[1], wc2[2], pch = 24, col = "black", bg = "white")


```

# Grabbing function
Smaller amounts of FW can be moved between adjacent ADs to equalize FW distribution. If a cluster is allocated FW between 100-139% of capacity, its FW points are reclassified to the least-filled adjacent AD. This in turn increases the amount of FW in the adjacent AD, minimizing the number of iterations of the FWADA model. 

## Sample clusters
In this example, the red cluster will be taking FW from the blue cluster.
```{r, echo = FALSE}

cluster2 <- km_house$BUS[km_house$BUS$AD_code == "x1",]

plot(km_house$BUS[km_house$BUS$AD_code %in% c("x1", "x7"),], pch = 20, col = "grey", main = "sample clusters | AD x1 and x7")
points(cluster, pch = 20, col = "red")
points(cluster2, pch = 20, col = "blue")
points(km_house$AD[km_house$AD$AD_code %in% c("x1", "x7"),], pch = 24, col = "black", bg = "white")


```

## FWd function
The FWd function is a variation of Ripley's K function. Instead of calculating number of points at a given distance from the center point, the FWd function calculates the amount of FW within a given distance from the center.

### Draw circles
Concentric circles are drawn around the AD that needs to increase its FW (the adjacent AD). The cumulative FW from the target AD (that needs to reduce its FW volume) within each circle is calculated. This cumulative FW is shown in the stepplot. The cutoff distance is determined by the capacity of the AD (red line).
```{r}

#FWd_func_one_point() - returns the cumulative FW as a function of distance from the a specified AD point.
# I basically just modified the one_point Kd function
#@param point - the point in question (usually lowest filled AD) that needs to "grab" FW
#@param other_BUS - the BUS points of the cluster being grabbed from
#@param radius - the radius of the FW_func and the breaks. Defaults to 50mi @ 1000 ft intervals
#returns - a df that can be plotted in ggplot
FWd_func_one_point <- function(point, other_BUS, radius = seq(1, 264000, 1000)) {
    

    # point distance bt lowest point and all others
    d <- raster::pointDistance(point, other_BUS, lonlat = FALSE)
    
    # a series of circles around events - every 100 feet from 0 to 15 miles
    distance2 <- radius
    
    # attach the pointdistance (d) to the good_NN df
    other_BUS$d_from_low <- d
    
    # 
    # for every distance, want the sum of FW for points within that distance
    FWd <- sapply(distance2, function(x) sum(other_BUS$FW_dis[other_BUS$d_from_low <= x]))
    
    FWd_df <- data.frame(kdist = distance2, FWd = FWd, num_ADs = nrow(other_BUS))
    
    return(FWd_df)
} # close FWd_one_point


FWd <- FWd_func_one_point(point = km_house$AD[km_house$AD$AD_code == "x7",], 
                          other_BUS = km_house$BUS[km_house$BUS$AD_code == "x7",],
                          radius = seq(1,10000, 500))


```

```{r}


plot(FWd$kdist, FWd$FWd, type = 's',
     main = "Cumulative FW by Distance | AD x7", 
     xlab = "Distance (feet)", 
     ylab = "Cumulative FW (tons/year)")
abline(h = 5000, col = "red")  
abline(v = max(FWd$kdist[FWd$FWd < 5000]), lty = 2)


plotCircle <- function(x, y, r) {
  angles <- seq(0,2*pi,length.out=360)
  lines(r*cos(angles)+x,r*sin(angles)+y)
}


plot(km_house$BUS[km_house$BUS$AD_code %in% c("x1", "x7"),], pch = 20, col = "grey", main = "draw circles")
points(cluster, pch = 20, col = "red")
points(cluster2, pch = 20, col = "blue")
points(km_house$AD[km_house$AD$AD_code %in% c("x1", "x7"),], pch = 24, col = "black", bg = "white")
lapply(seq(1,10000, 500), function(x) plotCircle(km_house$AD[km_house$AD$AD_code == "x7",]@coords[1],
                                                    km_house$AD[km_house$AD$AD_code == "x7",]@coords[2], x))




```

### Reclassify points
Points within this distance get reclassified to the new AD and FW in each cluster is calculated. The weighted center is used to determine new AD locations for each cluster (not shown in code here). In this example, there is very little of the blue cluster left after the grabbing function. Most likely, its FW will be distributed to other adjacent ADs in later iterations of the FWADA model and the blue cluster will be removed from the map.
```{r}

# add FW already in adjacent AD to the FWd df
FWd$FWd_plus <- FWd$FWd + km_house$AD$FW_dis[km_house$AD$AD_code == "x7"]


        
## change AD_code for BUS points associated with lowest_AD that are given to the of_adj_AD[it]

### find dist from overfilled_AD's BUS to of_adj_AD[it]
d <- raster::pointDistance(km_house$AD[km_house$AD$AD_code == "x7",],
                           km_house$BUS[km_house$BUS$AD_code == "x1",], lonlat = FALSE)

### attach this distance to the BUS df
km_house$BUS$pt_dist <- NA
km_house$BUS$pt_dist[km_house$BUS$AD_code == "x1"] <- d



### change AD_codes for points w/in FWd_xint distance to the of_adj_AD[it] code
### BUS where code is same as lowest AD and within the distance to add to the NN_AD
km_house$BUS$AD_code[km_house$BUS$AD_code == "x1" & km_house$BUS$pt_dist <= 5000] <- "x7"




# plot
plot(km_house$BUS[km_house$BUS$AD_code %in% c("x1", "x7"),], pch = 20, col = "grey", main = "grab FW")
points(km_house$BUS[km_house$BUS$AD_code == "x1",], pch = 20, col = "blue")
points(km_house$BUS[km_house$BUS$AD_code == "x7",], pch = 20, col = "red")
points(km_house$AD[km_house$AD$AD_code %in% c("x1", "x7"),], pch = 24, col = "black", bg = "white")
lapply(seq(1,10000, 500), function(x) plotCircle(km_house$AD[km_house$AD$AD_code == "x7",]@coords[1],
                                                    km_house$AD[km_house$AD$AD_code == "x7",]@coords[2], x))


```
